WIP: Transparent Standby for Low-Power, Resource-Constrained Embedded Systems

### 2018-02-20 LCTES/body.tex
Sugestões Noemi e Sztajnberg:   - colocar o modelo síncrono como um pré requisito   - incluir os desafios de software relacionados ao standby efetivo

According to the International Energy Agency (IEA), the number of
network-connected devices is expected to reach 50 billion by 2020 with the
expansion of the Internet of Things (IoT)~\cite{iea.data}.
%
However, most of the energy to power these devices will be consumed in
\emph{standby mode}, i.e., when they are neither transmitting or processing
data.
%
For instance, standby power currently accounts for approximately 10--15\% of
residential electricity consumption, and $CO_2$ emissions related to standby
are equivalent to those of 1 million cars~\cite{iea.data,standby.australia}.
%
The projected growth of IoT devices, together with the surprising effects of
standby consumption, made network standby efficiency one of the six
pillars of the \emph{G20's Energy Efficiency Action Plan}%
\footnote{G20's Energy Efficiency Action Plan: \url{https://www.iea-4e.org/projects/g20}}.

Given the projected scale of the IoT and the role of low-power standby towards
energy efficiency, our research has the following goals:

\begin{enumerate}
\item Address energy efficiency through extensive use of standby.
\item Target low-power, resource-constrained embedded architectures that form
      the IoT.
\item Provide standby mechanisms at the programming language level that scale
      to all applications.
\item Support transparent/non-intrusive standby mechanisms that reduce barriers
      of adoption.
\end{enumerate}

Our approach lies at the bottom of the software development
layers---programming language mechanisms---meaning that \emph{all} applications
should take advantage of low-power standby modes automatically, without extra
programming efforts.
%
We extend the synchronous programming language
\CEU~\cite{ceu.sensys13,ceu.tecs17} with support for interrupt service routines
(ISRs) and with a simple power management runtime (PMR).

@@

In Section~\ref{sec.ceu}, we compare the structure of programs in \CEU and
Arduino~\cite{arduino.book}, whose primary goal is to reduce the barrier of
adoption for a non-technical audience (e.g., designers and artists).
We show that we can keep the intended sequential reasoning of Arduino even when
applications require non-trivial concurrent behavior.
%
In Section~\ref{sec.standby}, we discuss the software infrastructure that
allows for unmmodified programs in \CEU to take advantage of standby
automatically.
%
In Section~\ref{sec.conclusion}, we discuss future work and conclude the paper.

>>>>

According to the International Energy Agency (IEA), the number of
network-connected devices is expected to reach 50 billion by 2020 with the
expansion of the Internet of Things (IoT)~\cite{iea.data}.
%
However, most of the energy to power these devices will be consumed in
\emph{standby mode}, i.e., when they are neither transmitting or processing
data.
%
For instance, standby power currently accounts for approximately 10--15\% of
residential electricity consumption, and $CO_2$ emissions related to standby
are equivalent to those of 1 million cars~\cite{iea.data,standby.australia}.
%
The projected growth of IoT devices, together with the surprising effects of
standby consumption, made network standby efficiency one of the six
pillars of the \emph{G20's Energy Efficiency Action Plan}%
\footnote{G20's Energy Efficiency Action Plan: \url{https://www.iea-4e.org/projects/g20}}.
%
However, making effective use of standby requires software-related efforts in
order to detect idle periods of activity in the device, identify peripherals
that must remain functional, and apply appropriate sleep mode levels in the
microcontroller.

Given the projected scale of the IoT, the role of low-power standby towards
energy efficiency, and the posed software-related challenges, our research has
the following goals:
%
%\begin{enumerate}
(i)   address energy efficiency through meticulous use of standby;
(ii)  target low-power, resource-constrained embedded architectures that form
      the IoT;
(iii) provide standby mechanisms at the programming language level that scale
      to all applications; and
(iv)  support transparent/non-intrusive standby mechanisms that reduce barriers
      of adoption.
%\end{enumerate}

Our approach lies at the bottom of the software development
layers---programming language mechanisms---meaning that \emph{all} applications
should take advantage of low-power standby modes automatically, without extra
programming efforts.
%
We extend the programming language \CEU~\cite{ceu.sensys13,ceu.tecs17} with
support for interrupt service routines (ISRs) and with a simple power
management runtime (PMR).
%
Our approach relies on the synchronous semantics of the language which enforces
that reactions to the environment always reach an idle state amenable to
standby.

@@

In Section~\ref{sec.ceu}, we compare the structure of programs in \CEU and
Arduino~\cite{arduino.book}, whose primary goal is to reduce the barrier of
adoption for a non-technical audience (e.g., designers and artists).
We show that we can keep the intended sequential reasoning of Arduino even when
applications require non-trivial concurrent behavior.
%
In Section~\ref{sec.standby}, we discuss the software infrastructure that
allows for unmodified programs in \CEU to take advantage of standby
automatically.
%
In Section~\ref{sec.conclusion}, we discuss future work and conclude the paper.

### 2018-02-20 LCTES/lctes.tex
Sugestões Noemi e Sztajnberg:   - colocar o modelo síncrono como um pré requisito   - incluir os desafios de software relacionados ao standby efetivo

\begin{abstract}
Standby efficiency for connected devices is one of the priorities of the
\emph{G20's Energy Efficiency Action Plan}.
%
We propose transparent programming language mechanisms to enforce that
applications remain in the deepest standby modes for the longest periods of
time.
%
We extend the synchronous programming language \CEU with support for interrupt
service routines and with a simple power management runtime.
%
We provide device drivers based on these primitives on top of which
applications can be built to take advantage of standby automatically.
%
We also show that programs in \CEU can keep a sequential structure to lower the
barrier of adoption, even when applications require non-trivial concurrent
behavior.
\end{abstract}

>>>>

\begin{abstract}
Standby efficiency for connected devices is one of the priorities of the
\emph{G20's Energy Efficiency Action Plan}.
%
We propose transparent programming language mechanisms to enforce that
applications remain in the deepest standby modes for the longest periods of
time.
%
We extend the programming language \CEU with support for interrupt service
routines and with a simple power management runtime.
%
We also provide device drivers based on these primitives on top of which
applications can be built to take advantage of standby automatically.
%
Our approach relies on the synchronous semantics of the language which enforces
that reactions to the environment always reach an idle state amenable to
standby.
%
In addition, we show that programs in \CEU can keep a sequential structure,
even when applications require non-trivial concurrent behavior, to lower the
barrier of adoption.
\end{abstract}
